%%% Appendices — content only.
%%% Each \chapter inside this file becomes an appendix (A, B, C, ...),
%%% because main.tex sets \multipleAppendixtrue. The class hijacks
%%% the \chapter command in \AtEndDocument to step an appendix counter
%%% and label entries APPENDIX A, APPENDIX B, etc., in the ToC.

\chapter{SUPPLEMENTARY VALIDATOR RULES}\label{app:rules}

This appendix lists rule identifiers and citations that did not warrant
inline discussion in the main chapters but are checked by the
validator. See \texttt{docs/uf-rules.md} for full detection strategies
and source citations.

\begin{itemize}
    \item UF-D1 -- editMode option present
    \item UF-D2 -- LuaLaTeX directive
    \item UF-D3 -- overrideTitles / overrideChapters options
    \item UF-P1 -- \texttt{\textbackslash set*File} companions exist on disk
\end{itemize}

\chapter{SAMPLE BROKEN INPUTS}\label{app:fixtures}

To verify that each must-fix rule fires, the test suite includes
synthetic broken inputs that violate exactly one rule. Each lives in
\texttt{tests/fixtures/broken/<rule-id>/} and includes a brief note
explaining what was broken and why the corresponding rule should fire.

A clean release of the validator detects every broken fixture and
flags it with the corresponding rule identifier.

\chapter{SUBMISSION CHECKLIST FOR THE STUDENT}\label{app:checklist}

The validator catches mechanical formatting violations. It does not
catch everything. Before clicking submit, work through this checklist.
Your committee, your department, and the UF Editorial Office will
appreciate every box you mark.

\section{Things the validator checks for you}

\begin{itemize}
    \item Margins, fonts, line spacing, alignment, paragraph indent
          (rules UF-F1 through UF-F7).
    \item Page numbering scheme (UF-F6).
    \item Required sections present and in order (UF-F8, UF-F9).
    \item At least three chapters (UF-F10).
    \item Required metadata macros set (UF-F14).
    \item Abstract under the 350-word cap (UF-F15).
    \item Broken internal references and uncited bibliography entries
          (UF-S3).
    \item Use of LuaLaTeX directive (UF-D2).
\end{itemize}

\section{Things only you can check}

\begin{itemize}
    \item Your committee has signed the ETD signature page in GIMS.
    \item Your degree application is current in ONE.UF.
    \item Your dissertation title in the title page matches the title
          on your degree application.
    \item Your major matches your department's official major name.
    \item Your committee chair's name is spelled correctly. (And the
          co-chair's, if applicable.)
    \item The graduation month is May, August, or December and matches
          your intended graduation term.
    \item Every figure has a caption. Every table has a caption. Every
          equation that you cite has a label.
    \item Every citation in your text resolves to a real entry in
          your bibliography, and every bibliography entry is cited at
          least once.
    \item Your acknowledgments thank the people who actually got you
          here. (They will read it.)
    \item Co-authored chapters acknowledge co-authors.
    \item Published-chapter footnotes appear on the first page of each
          published-chapter, with the journal citation.
    \item Image resolution is high enough to read on screen.
    \item Hyperlinks in your PDF actually go where you expect.
    \item You have a backup of the source.
    \item You have a backup of the backup.
    \item You have eaten today.
    \item You have slept in the last twenty-four hours, ideally more
          than four.
\end{itemize}

\section{One last note}

You wrote a dissertation. Whatever happens at submission, that fact
does not go away. Take a breath, run the checklist, and submit.

You earned this.
