%%% Biographical sketch — content only.
%%% The class inserts the \chapter{BIOGRAPHICAL SKETCH} header
%%% automatically. Third-person prose is conventional. UF does not
%%% require this to be solemn.

Jordan Avery Lee discovered LaTeX in their first semester of graduate
school and discovered the UF formatting specification at 11:43~p.m.\ the
night before their First Submission deadline. Both experiences shaped
their dissertation.

Before graduate school, Lee earned a Bachelor of Science in Computer
Science from a public research university whose graduate office,
charitably, used Microsoft Word templates. Lee entered the doctoral
program at the University of Florida the same year and devoted the
next several years to research, teaching, two failed thesis title
revisions, one successful committee reshuffle, and the cultivation of
strong opinions about page breaks. They served as a graduate teaching
assistant for three semesters and as a graduate research assistant for
the remainder of their training, accruing along the way an unreasonable
familiarity with the \texttt{titlesec} package.

Their research interests include static analysis, software tooling for
academic workflows, and accessibility in technical documents. Their
non-research interests include long walks, longer compile times, and
the persistent suspicion that one more pass of \texttt{lualatex} will
resolve the cross-reference warnings. (It usually does.)

Lee received the doctoral degree in May 2026. The dissertation
compiled on the first try.\footnote{This is a lie; it compiled on the
fourth try. The first three produced increasingly creative undefined
references, but the dissertation is, after all, finished.}
